### 2022-02-28 06_Paper/ER-ED/method.Rmd
Update method.Rmd
Update Analysis plan

---
output: html_document
bibliography: ER_ED.bib
csl: apa7.csl
editor_options: 
  markdown: 
    wrap: sentence
---

```{r SetupMethod, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(here)       # to set directories without defined paths
# set top level directory to source file
here::i_am("flag_project_root_CERED.txt")

library(papaja)     # APA conform output
library(readxl)     # import excel files
```

```{r stimvalues, echo = FALSE}
# import excel list with neutral and negative pictures
# lists contain arousal and valence values of each picture
stim_negative <- readxl::read_xlsx(file.path(here("01_Paradigms", "ER-ED", "ER_ED", "Stimuli", "negative.xlsx")), range = cell_cols("A:D"))

stim_neutral <- readxl::read_xlsx(file.path(here("01_Paradigms", "ER-ED", "ER_ED", "Stimuli", "neutral.xlsx")), range = cell_rows(1:21))
```

# 2. Method

We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study [@Simmons2012].
The paradigm was written and presented using *Psychopy* [@Peirce2019].
We used *R* with *R Studio* [@RCore2021; @RStudioTeam2020] with the main packages *afex* [@Singmann2021] and *BayesFactor*[@Morey2021] for all analyses.
The R Markdown file used to analyze the data and write this document, as well as the raw data and the materials are freely available at https://github.com/ChScheffel/CERED.<!-- eventually perhaps refer to the _script and R Markdown file_ -->
A complete list of all measures assessed in the study can be found at OSF (<https://osf.io/vnj8x/>) and GitHub (https://github.com/ChScheffel/CERED).
<!-- Attention! OSF is still private -->

## 2.1 Ethics information

The study protocol complies with all relevant ethical regulations and was approved by the ethics committee of the Technische Universität Dresden (reference number EK...). <!-- Attention!add EK number, when approved --> 
Prior to testing, written informed consent will be obtained.
Participants will receive 30€ in total or course credit for participation.

## 2.2 Pilot data

The procedure described above was tested in a pilot study with $N=`r nrow(data_quest_pilot)`$ participants (`r length(which(data_quest_pilot$gender == 2))` female; age: $M = `r papaja::printnum(mean(data_quest_pilot[,"age"], na.rm = TRUE), digits = 1, gt1 = TRUE)`\text{ }\pm\text{ }SD = `r papaja::printnum(sd(data_quest_pilot[,"age"], na.rm = TRUE), digits = 1, gt1 = TRUE)`$).
<!-- direkt "above" wird das nicht beschrieben, die genaue procedure wird eigentlich auch erst unten beschrieben. Ich würde eher schreiben "The newly developed ER paradigm was tested in a pilot study" oder so ähnlich. --> 
Regarding self-reported arousal, results showed significant higher subjective arousal for active viewing of negative compared to neutral pictures.
<!-- Signifikanz beruht auf p-Werten? Würde ich angeben. --> 
However, ER strategies did not lead to a reduction of subjective arousal compared to active viewing of negative pictures.
Regarding physiological responses, ER strategies were associated with reduced facial muscle activity of the *corrugator* and *levator* compared to active viewing of negative pictures.
In accordance with our previous study [@Scheffel2021], we found that the use of ER strategies compared to active viewing was associated with increased subjective effort.
<!-- kann man die Results irgendwo nachvollziehen? Ich würde zumindest MW, SD und p-Werte angeben wenn gesagt wird, dass es "signifikant" höher/nierdiger ist. Auch um zu zeigen, worauf sich die Aussage stützt. 
könnte man die Ergebnisse sogar in einer Tabelle/Abbildung darstellen? Je nachdem, wie wichtig die Pilotdaten für das Gesamt-Ms. sind...

Ich war etwas verwirrt dass die Pilotdaten schon hier kommen. Ich hätte erwartet, dass erst das Paradigma genau beschrieben wird, und dann als "Schmankerl" die ersten Pilotdaten mitgeliefert werden. Evtl. mit einer extra Überschrift "Introduction / Methods / Pilot study (oder data)". So gehen sie m.E. etwas unter, weil man sich nachfolgend wieder auf die Details des Paradigmas konzentriert. Hier wurde aber auch nicht alles untersucht, wie z.B. Fragebögen? -->

## 2.3 Design

Young healthy participants (aged 18 to 30 years) will be recruited using the software *ORSEE* [@Greiner2015] at the Technische Universität Dresden.
Participants will be excluded from participation if they do not fluently speak German, have current or a history of psychological disorders or neurological trauma, or report to take medication.
Participants will be invited to complete an online survey containing different questionnaires to assess broad and narrow personality traits and measures of well-being.
The study consists of two lab sessions, which will take place in a shielded cabin with constant lighting.
Before each session, participants will receive information about the respective experimental procedure and provide informed consent.
At the beginning of the first session participants will fill out a demographic questionnaire and complete an n-back task with the levels one to four.
Then, they will complete an effort discounting (ED) procedure on screen, followed by a random repetition of one n-back level.
The second session will comprise the ER paradigm and take place exactly one week after session one.
Participants will provide informed consent and receive written instructions on the ER paradigm and ER strategies that they should apply.
A brief training will ensure that all participants are able to implement the ER strategies.
Next, electrodes to measure facial EMG will be attached and the ER paradigm will be conducted.
Participants will receive 30.00€ or course credit as compensation.
<!-- das steht schon bei Ethics information, evtl. kürzen. -->
Study data will be collected and managed using REDCap electronic data capture tools hosted at Technische Universität Dresden [@Harris2009; @Harris2019].

### 2.3.1 Psychometric measures

The online survey contains a number of questionnaires: General psychological well-being is assessed using the German version of the WHO-5 scale [@Bech2004; @Braehler2007].
<!-- Hier und im Folgenden wird wieder Präsens verwendet - vereinheitlichen? -->
To measure resilience, the German version 10-item-form of the Connor-Davidson resilience Scale (CD-RISC)[@Connor2003,@Campbell-Sills2007; @Sarubin2015] is used.
Habitual use of ER is assessed using the German version of the Emotion Regulation Questionnaire (ERQ) [@GrossJohn2003; @Abler2009].
For the assessment of ER flexibility we use the Flexible Emotion Regulation Scale (FlexER) [@Doerfel2019].
Implicit theories of willpower in emotion control are assessed using the implicit theories questionnaire from @Bernecker2017.
To assess Need for Cognition, the German version short form of the Need for Cognition Scale [@Cacioppo1982; @Bless1994] is used.
To assess self-control, sum scores of the German versions of the following questionnaires are used [@Paschke2016]: the Self-Regulation Scale (SRS) [@Schwarzer1999], the Brief Self-Control Scale (BSCS) [@Tangney2004; @Sproesser2011], and the Barratt Impulsiveness Scale (BIS-11) [@Patton1995; @Hartmann2011].
Attentional control is assessed using the Attentional Control Scale (ACS) [@Derryberry2002].
For more detailed information on psychometric properties of the questionnaires, please see [supplementary material](#SupplementQuestionnaires). <!-- Not sure whether you can state this so briefly without giving more detailed information on, e.g., sample items, the scale on which answers are recorded (e.g., -3 ... +3), the reliability of the measures. 
AG: im Paper würde man das hier angeben, bei RR reicht vielleicht Supplememt? -->

### 2.3.2 Emotion regulation paradigm

The ER paradigm consists of three parts that will be described in the following.

*Part one: ER task.* <br> Part one is a standard ER task in a block design (see Figure X), similar to paradigms previously used by our group [@Scheffel2021].
Participants are told to actively view neutral and negative pictures (see [2.3.3](##stimuli)) or to regulate all upcoming emotions by means of distraction, distancing, and expressive suppression, respectively.
Every participant first has the condition "active viewing-neutral" that serves as a baseline condition.
During this block, 20 neutral pictures are presented.
Participants are asked to "actively view all pictures and permit all emotions that may arise." In the second block, participants actively viewe negative pictures.
During the third, fourth, and fifth block, participants see negative pictures and are asked to regulate their emotions using distraction, distancing, and suppression.
In order to achieve distraction, participants are asked to think of a geometric object or an everyday activity, like brushing their teeth.
During distancing, participants are asked to "take the position of a non-involved observer, thinking about the picture in a neutral way." Participants are told not to re-interpret the situation or attaching a different meaning to the situation.
During suppression, participants are told to "suppress their emotional facial expression." They should imagine being observed by a third person that should not be able to tell by looking at the facial expression whether the person is looking at an emotional picture.
Participants are instructed not to suppress their thoughts or change their facial expression to the opposite [@Scheffel2021].
All participants receive written instruction and complete a training session.
After the training session, participants are asked about their applied ER strategies to avoid misapplication.
The order of the three regulation blocks (distraction, distancing, and suppression) is randomized between participants.

*Part two: ER effort discounting.* <br> In the second part, ER effort discounting takes place.
The procedure of the discounting follows the COG-ED paradigm by Westbrook et al. [-@Westbrook2013] with a major change.
We use the following adaption that allows the computation of SVs for different strategies without presuming that all individuals would inherently evaluate the same strategy as the easiest one: For each possible pairing (distraction vs. distancing, distraction vs. suppression, and distancing vs. suppression), two strategies with monetary values are presented.
The order of the comparisons is randomized.
Because there is no strategy that is objectively more difficult, we added an initial comparison that begins with the option "1 EUR for strategy A or 1 EUR for strategy B".
The strategy that is not chosen is assigned the value of 2 EUR.
From this point on, comparisons between strategies follow the original COG-ED paradigm [@Westbrook2013].
Participants are instructed to decide as realistically as possible by imagining that the money displayed is actually available for choice.

*Part three: ER choice.* <br> After the discounting part, participants choose which of the three ER strategies (distraction, distancing or suppression) they want to re-apply.
Importantly, there is no further instruction on what basis they should make their decision.
Participants should make their decision freely, according to criteria they consider important for themselves.
However, participants are asked to state the reasons for the decision afterwards.
As soon as they have decided, they see the respective instruction and the block with another 20 negative pictures starts.

### 2.3.3 Stimuli

Pictures used in the paradigm were selected from the Emotional Picture Set (EmoPicS) [@Wessa2010] and the International Affective Picture System (IAPS) [@Lang2008].
The 20 neutral pictures (Valence (V): *M* $\pm$ *SD* = `r round(mean(stim_neutral$valence), digits = 2)` $\pm$ `r round(sd(stim_neutral$valence), digits = 2)`; Arousal (A): *M* $\pm$ *SD* = `r round(mean(stim_neutral$arousal), digits = 2)` $\pm$ `r round(sd(stim_neutral$arousal), digits = 2)`) depicted content related to the categories persons, objects, and scenes.
Further, 100 negative pictures, featuring categories animals, body, disaster, disgust, injury, suffering, violence, and weapons, were used.
An evolutionary algorithm [@Yu2010] was used to cluster these pictures into five sets with comparable valence and arousal values (set one: V: *M* $\pm$ *SD* = `r round(mean(stim_negative$valence[stim_negative$set == 1]), digits = 2)` $\pm$ `r round(sd(stim_negative$valence[stim_negative$set == 1]), digits = 2)`, A: *M* $\pm$ *SD* = `r round(mean(stim_negative$arousal[stim_negative$set == 1]), digits = 2)` $\pm$ `r round(sd(stim_negative$arousal[stim_negative$set == 1]), digits = 2)`; set two: V: *M* $\pm$ *SD* = `r round(mean(stim_negative$valence[stim_negative$set == 2]), digits = 2)` $\pm$ `r round(sd(stim_negative$valence[stim_negative$set == 2]), digits = 2)`, A: *M* $\pm$ *SD* = `r round(mean(stim_negative$arousal[stim_negative$set == 2]), digits = 2)` $\pm$ `r round(sd(stim_negative$arousal[stim_negative$set == 2]), digits = 2)`; set three: V: *M* $\pm$ *SD* = `r round(mean(stim_negative$valence[stim_negative$set == 3]), digits = 2)` $\pm$ `r round(sd(stim_negative$valence[stim_negative$set == 3]), digits = 2)`, A: *M* $\pm$ *SD* = `r round(mean(stim_negative$arousal[stim_negative$set == 3]), digits = 2)` $\pm$ `r round(sd(stim_negative$arousal[stim_negative$set == 3]), digits = 2)`; set four: V: *M* $\pm$ *SD* = `r round(mean(stim_negative$valence[stim_negative$set == 4]), digits = 2)` $\pm$ `r round(sd(stim_negative$valence[stim_negative$set == 4]), digits = 2)`, A: *M* $\pm$ *SD* = `r round(mean(stim_negative$arousal[stim_negative$set == 4]), digits = 2)` $\pm$ `r round(sd(stim_negative$arousal[stim_negative$set == 4]), digits = 2)`; set five: V: *M* $\pm$ *SD* = `r round(mean(stim_negative$valence[stim_negative$set == 5]), digits = 2)` $\pm$ `r round(sd(stim_negative$valence[stim_negative$set == 5]), digits = 2)`, A: *M* $\pm$ *SD* = `r round(mean(stim_negative$arousal[stim_negative$set == 5]), digits = 2)` $\pm$ `r round(sd(stim_negative$arousal[stim_negative$set == 5]), digits = 2)`).
A complete list of all pictures and their classification into sets can be found in supplementary material 1.

### 2.3.4 Facial electromyography

Bipolar facial electromyography (EMG) will be measured for *corrugator supercilii* and *levator labii* as indices of affective valence [@Bradley2000], similar to previous work by our group [@Gärtner, Jawinski, Strobel, 2021].
<!-- da wir das ja sogar schon gemacht haben, fände ich es cool die Studie zu erwähnen. :-) Zeigt ja auch, dass wir keine Neulinge auf dem Gebiet sind. Referenz muss glaub noch in Bib eingefügt werden! --> 
Two passive surface Ag/AgCl electrodes (8 mm inner diameter, 10 mm distance between electrodes) will be placed over each left muscle according to the guidelines of @Fridlund1986.
The ground electrode will be placed over the left *Mastoid*.
Before electrode placement, the skin will be abraded with Every abrasive paste, cleaned with alcohol, and filled with Lectron III electrolyte gel.
Raw signals will be amplified by a BrainAmp amplifier (Brain Products Inc., Gilching, Germany).
Impedance level will be kept below 10 $k\Omega$. 
Data will be sampled at 1000 Hz, filtered, rectified and integrated. 
A 20 Hz high pass (order 8), a 300 Hz low pass (order 8), and a 50 Hz notch filter will be applied to both signals.
Corrugator and Levator EMG will be analyzed during the 6 s of picture presentation.
EMG data will be baseline-corrected using a time window of 2 s prior to stimulus onset [@Bradley2000]. 
Last, the sampling rate will be changed to 100 Hz, and EMG data will be averaged for each condition and each participant.

## 2.4 Sampling plan

Sample size calculation is done using *G\*Power* [@Faul2007; @Faul2009].
In a meta-analysis of Zaehringer and colleagues [-@Zaehringer2020], effect sizes of ER on peripheral-physiological measures were reported:
To find an effect of $d=-0.32$ of ER on *corrugator* muscle activity with $\alpha=.05$ and $\beta=.95$, data of at least $N=85$ have to be analyzed.
Power analyses of all other hypotheses yielded smaller sample sizes.
However, if participants withdraw from study participation, technical failures occur, or experimenter considers the participant for not suitable for study participation (e.g., because the participant does not follow instructions or shows great fatigue), respective data will also be excluded from further analyses.
Therefore, we aim to collect data of 90 participants.
Detailed information on power calculation for each hypothesis can be found in the [design table](#DesignTable) in the Supplement.
<!-- hier würde ich den Verweis zur Design Table ganz explizit machen, also sagen, wo diese zu finden ist. -->

## 2.5 Analysis plan
<!-- hier wird wieder durchgehend Futur verwendet - würde das in Methods vereinheitlichen. -->

Data collection and analysis will not be performed blind to the conditions of the experiments.
The level of significance will be set to $\alpha=.05$.

<!-- Bei den krusiven Zwischenüberschriften wie in der nächsten Zeile würde ich Punkte setzen, sieht sonst im PDF sehr eng zusammen aus. ;-)  -->
*Effects of emotion regulation on arousal, facial EMG and effort.* <br> To examine the impact of valence of emotional pictures on subjective arousal ratings, a repeated measures analysis of variance (rmANOVA) with the factor valence (neutral and negative) for the strategy active viewing will be conducted.
To investigate the effects of the three ER strategies on subjective arousal, another rmANOVA  with the factor strategy (active viewing - negative, distraction, distancing, and suppression) for subjective arousal ratings will be conducted.

To examine the impact of valence on facial EMG, a rmANOVA with the factor valence (neutral and negative) for the strategy active viewing will be conducted for *corrugator* and *levator* activity.
To examine the effects of the three ER strategies on facial EMG, another rmANOVA with the factor strategy (active viewing - negative, distraction, distancing, and suppression) for *corrugator* and *levator* activity will be conducted.

To examine the effect of ER strategies on cognitive effort, a rmANOVA with the factor strategy (active viewing, distraction, distancing, and suppression) for subjective effort ratings will be conducted.
Greenhouse-Geisser-corrected $p$-values and degrees of freedom will be reported when the assumption of sphericity is violated.
Proportion of explained variance $\eta_{p}^{2}$ will be reported as a measure of effect size.
If indicated by the data, estimated marginal means will be computed as post-hoc contrasts.
rmANOVAS will be run using the *afex* package [@Singmann2021]

*Subjective values of emotion regulation strategies.* <br> For each ER strategy, SVs will be calculated as follows: first, the value 0.015625 will be added to or subtracted from the last monetary value of the flexible strategy, depending on the participant's last choice.
Second, the resulting (monetary) value will be divided by 2.00 €.
The final SV for each participant will be computed by averaging all final SVs of each strategy.
The resulting values will be entered in a rmANOVA to compare the SVs of the three strategies (distraction, distancing, and suppression) and explore group effects.
Again, estimated marginal means will be computed as post-hoc contrasts.

To investigate whether individual SVs predict ER choice, a Chi-squared test with predicted choice (highest SV of each participant) and actual choice will be computed.
Furthermore, an ordinal logistic regression with the dependent variable choice and independent variables SVs of each strategy will be computed.

To explore the association between subjective arousal, physiological responses, and subjective effort on SVs, a multilevel model (MLM) will be specified using the $lmerTest$ package [@Kuznetsova2017].
First, ER strategies will be recoded and centered for each subject according to their individual SVs: The strategy with the highest SV will be coded as -1, the strategy with the second highest SV 0, and the strategy with the lowest SV will be coded as 1.
Restricted maximum likelihood (REML) will be applied to fit the model.
A random slopes model of SVs including subjective effort (effort ratings), subjective arousal (arousal ratings), and physiological responses (*corrugator* and *levator* activity) as level-1-predictors will be specified.
$$
SV \sim strategy\ + \text{effort rating} + \text{arousal rating} + corrugator \text{ activity} + levator \text{ activity} + (strategy|subject)
$$ Level-1-predictors will be centered within cluster [@Enders2007].
Residuals of the final model will be inspected visually.
Intraclass correlation coefficient (ICC), $\rho$, will be reported for each model (null model, as well as full model).

The influence of personality traits on SVs will be investigated exploratorily.
Therefore, the MLM specified above will be extended by the level-2-predictors NFC and self-control.

The association between flexible ER and SVs of ER strategies will be investigated with a regression using the $intercept$ and $slope$ of each participants' SVs to predict their FlexER score.
Firstly, SVs will be ordered by magnitude.
Secondly, for each participant a linear model will be built to estimate the individual $intercept$ and $slope$.

For each result of the analyses, both $p$-values and Bayes factors $BF10$, calculated using the *BayesFactor* package [@Morey2021], will be reported. Bayes factors are calculated using the default prior widths of the functions *anovaBF*, *lmBF* and *regressionBF*.

## Data availability

The data of this study can be downloaded from [osf.io/vnj8x/](https://osf.io/vnj8x/).

## Code availability

The paradigm code, as well as the R Markdown file used to analyze the data and write this document is available at our [Github repository](https://github.com/ChScheffel/CERED).