### 2023-08-31 06_Paper/COG-ED/Stage 2/Supplement.Rmd

---
title             : "Supplementary Material for the Registered Report titled: 'When easy is not preferred: A discounting paradigm to assess load-independent task preference'"
shorttitle        : "Supplement for CAD Paradigm"
author: 
  - name          : "Josephine Zerna"
    orcid         : "0000-0003-2892-884X"
    affiliation   : "1"
    corresponding : yes
    address       : "Zellescher Weg 17, 01069 Dresden, Germany"
    email         : "josephine.zerna@tu-dresden.de"
    equal_contrib : yes
    role:         # Contributorship roles (e.g., CRediT, credit.niso.org)
      - Conceptualization
      - Data curation
      - Methodology
      - Funding acquisition
      - Formal analysis
      - Investigation
      - Project administration
      - Software
      - Visualization
      - Writing - original draft
      - Writing - review & editing
  - name          : "Christoph Scheffel"
    orcid         : "0000-0001-5963-9229"
    affiliation   : "1"
    equal_contrib : yes
    role:
      - Conceptualization
      - Methodology
      - Funding acquisition
      - Investigation
      - Project administration
      - Software
      - Writing - review & editing
  - name          : "Corinna Kührt"
    orcid         : "0000-0002-6418-6479"
    affiliation   : "1"
    role:
      - Formal analysis
      - Writing - review & editing
  - name          : "Georgia Clay"
    orcid         : "0000-0001-5641-6804"
    affiliation   : "2"
    role:
      - Formal analysis
  - name          : "Alexander Strobel"
    orcid         : "0000-0002-9426-5397"
    affiliation   : "1"
    role: 
      - Conceptualization
      - Resources
      - Supervision
      - Funding acquistion
      - Writing - review & editing
affiliation:
  - id            : "1"
    institution   : "Faculty of Psychology, Technische Universität Dresden, 01069 Dresden, Germany"
  - id            : "2"
    institution   : "Faculty of Psychology, University of Vienna, Vienna 1010, Austria"
keywords          : "effort discounting, registered report, specification curve analysis, need for cognition, $n$-back"
bibliography      : CAD.bib
floatsintext      : yes
figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no
documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
csl               : nature.csl
always_allow_html : true
header-includes   :
    - \usepackage{booktabs}
    - \usepackage{longtable}
    - \usepackage{array}
    - \usepackage{multirow}
    - \usepackage{wrapfig}
    - \usepackage{float}
    - \usepackage{colortbl}
    - \usepackage{pdflscape}
    - \usepackage{tabu}
    - \usepackage{threeparttable}
    - \usepackage{threeparttablex}
    - \usepackage[normalem]{ulem}
    - \usepackage{makecell}
    - \usepackage{xcolor}
    - \usepackage{setspace}\doublespacing
    - \usepackage[final]{pdfpages}
    - \usepackage{chngcntr}

---
\newpage
\renewcommand\thesection{\Alph{section}}
\counterwithin{figure}{section}
\counterwithin{table}{section}
\setcounter{section}{19}
\setcounter{figure}{0}
\setcounter{table}{0}

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}

  # set the output for figures and tables to the width of the text throughout the entire script

  knitr::opts_chunk$set(out.width = "\\textwidth")

  # load libraries

  library(bibtex)       # for referencing, "bibtex" package
  library(here)         # for path independent file referencing, "here" package
  library(papaja)       # APA conform output, "papaja" package
  options(papaja.na_string = "")
  library(kableExtra)   # for great tables in R Markdown, "kableExtra" package
  library(interactions) # for simple slopes plots, "interactions" package
  library(BayesFactor)  # computing Bayes factors, "BayesFactor" package
  library(ggplot2)      # for plotting, "ggplot2" package
  library(knitr)        # for including plots and such, "knitr" package
  library(effectsize)   # for effect sizes and CIs, "effectsize" package

  # set top level directory to source file

  here::i_am("flag_project_root_CAD.txt")

  # load the workspace variables that have been computed with the Analysis.R script

  load(here("06_Paper","COG-ED","Stage 2","Workspace.RData"))

```

## Design Table

(Starts on next page)

\includepdf[pages={-},landscape=true]{Design_Table_T1.pdf}

\newpage

## Contrasts tested for in the repeated measures ANOVA

```{r contrast-plot, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Contrasts tested for in the repeated measures ANOVA that was used as a prerequisite for the multi level model of hypothesis 2b. Ascending quadratic (-1,1,1,-1), declining linear (3,1,-1,-3), declining logistic (3,2,-2,-3), and positively skewed normal (1,2,-1,-2). Figure available at osf.io/vnj8x/, under a CC-BY-4.0 license.", fig.pos="H"}

  include_graphics(here("06_Paper", "COG-ED", "Stage 2", "Figures", "contrasts.png"))

```

\newpage

## Hypothesis 1a: The signal detection measure $d'$ declines with increasing $n$-back level.

ANOVA:

`r apa_print(hypothesis1a_rmanova)$full_result`, `r apa_print(hypothesis1a_BF)$statistic`

Paired contrasts:

```{r, echo=FALSE, message=FALSE, warning=FALSE}
apa_table(
  hypothesis1a_contrasts,
  caption = "Paired contrasts for the rmANOVA comparing $d'$ between $n$-back levels.",
  note = "The column Contrast contains the $n$ of the $n$-back levels. $SE$ = standard error, $df$ = degrees of freedom, $t$ = $t$-statistic, $p$ = $p$-value, CI = confidence interval.",
  escape = FALSE,
  placement = "H"
)
```

\newpage

## Hypothesis 1b: The median reaction time increases with increasing $n$-back level.

ANOVA:

`r apa_print(hypothesis1b_rmanova)$full_result`, `r apa_print(hypothesis1b_BF)$statistic`

Paired contrasts:

```{r, echo=FALSE, message=FALSE, warning=FALSE}
apa_table(
  hypothesis1b_contrasts,
  caption = "Paired contrasts for the rmANOVA comparing the median reaction time between $n$-back levels.",
  note = "The column Contrast contains the $n$ of the $n$-back levels. $SE$ = standard error, $df$ = degrees of freedom, $t$ = $t$-statistic, $p$ = $p$-value, CI = confidence interval.",
  escape = FALSE,
  placement = "H"
)
```

\newpage

## Hypothesis 1c: Ratings on all NASA-TLX dimensions increase with increasing $n$-back level.

Mental subscale ANOVA:

`r apa_print(hypothesis1c_mental_rmanova)$full_result`, `r apa_print(hypothesis1c_mental_BF)$statistic`

Mental subscale paired contrasts:

```{r, echo=FALSE, message=FALSE, warning=FALSE} 
apa_table(
  hypothesis1c_mental_contrasts,
  caption = "Paired contrasts for the rmANOVA comparing ratings on the NASA-TLX Mental subscale between $n$-back levels.",
  note = "The column Contrast contains the $n$ of the $n$-back levels. $SE$ = standard error, $df$ = degrees of freedom, $t$ = $t$-statistic, $p$ = $p$-value, CI = confidence interval.",
  escape = FALSE,
  placement = "H"
)
```

\newpage

Physical subscale ANOVA:

`r apa_print(hypothesis1c_physical_rmanova)$full_result`, `r apa_print(hypothesis1c_physical_BF)$statistic`

Physical subscale paired contrasts:

```{r, echo=FALSE, message=FALSE, warning=FALSE} 
apa_table(
  hypothesis1c_physical_contrasts,
  caption = "Paired contrasts for the rmANOVA comparing ratings on the NASA-TLX Physical subscale between $n$-back levels.",
  note = "The column Contrast contains the $n$ of the $n$-back levels. $SE$ = standard error, $df$ = degrees of freedom, $t$ = $t$-statistic, $p$ = $p$-value, CI = confidence interval.",
  escape = FALSE,
  placement = "H"
)
```

\newpage

Time subscale ANOVA:

`r apa_print(hypothesis1c_time_rmanova)$full_result`, `r apa_print(hypothesis1c_time_BF)$statistic`

Time subscale paired contrasts:

```{r, echo=FALSE, message=FALSE, warning=FALSE} 
apa_table(
  hypothesis1c_time_contrasts,
  caption = "Paired contrasts for the rmANOVA comparing ratings on the NASA-TLX Time subscale between $n$-back levels.",
  note = "The column Contrast contains the $n$ of the $n$-back levels. $SE$ = standard error, $df$ = degrees of freedom, $t$ = $t$-statistic, $p$ = $p$-value, CI = confidence interval.",
  escape = FALSE,
  placement = "H"
)
```

\newpage

Performance subscale ANOVA:

`r apa_print(hypothesis1c_performance_rmanova)$full_result`, `r apa_print(hypothesis1c_performance_BF)$statistic`

Performance subscale paired contrasts:

```{r, echo=FALSE, message=FALSE, warning=FALSE} 
apa_table(
  hypothesis1c_performance_contrasts,
  caption = "Paired contrasts for the rmANOVA comparing ratings on the NASA-TLX Performance subscale between $n$-back levels.",
  note = "The column Contrast contains the $n$ of the $n$-back levels. $SE$ = standard error, $df$ = degrees of freedom, $t$ = $t$-statistic, $p$ = $p$-value, CI = confidence interval.",
  escape = FALSE,
  placement = "H"
)
```

\newpage

Effort subscale ANOVA:

`r apa_print(hypothesis1c_effort_rmanova)$full_result`, `r apa_print(hypothesis1c_effort_BF)$statistic`

Effort subscale paired contrasts:

```{r, echo=FALSE, message=FALSE, warning=FALSE} 
apa_table(
  hypothesis1c_effort_contrasts,
  caption = "Paired contrasts for the rmANOVA comparing ratings on the NASA-TLX Effort subscale between $n$-back levels.",
  note = "The column Contrast contains the $n$ of the $n$-back levels. $SE$ = standard error, $df$ = degrees of freedom, $t$ = $t$-statistic, $p$ = $p$-value, CI = confidence interval.",
  escape = FALSE,
  placement = "H"
)
```

\newpage

Frustration subscale ANOVA:

`r apa_print(hypothesis1c_frustration_rmanova)$full_result`, `r apa_print(hypothesis1c_frustration_BF)$statistic`

Frustration subscale paired contrasts:

```{r, echo=FALSE, message=FALSE, warning=FALSE} 
apa_table(
  hypothesis1c_frustration_contrasts,
  caption = "Paired contrasts for the rmANOVA comparing ratings on the NASA-TLX Frustration subscale between $n$-back levels.",
  note = "The column Contrast contains the $n$ of the $n$-back levels. $SE$ = standard error, $df$ = degrees of freedom, $t$ = $t$-statistic, $p$ = $p$-value, CI = confidence interval.",
  escape = FALSE,
  placement = "H"
)
```

\newpage

## Hypothesis 2a: Subjective values decline with increasing $n$-back level.

ANOVA:

`r apa_print(hypothesis2a_rmanova)$full_result`, `r apa_print(hypothesis2a_BF)$statistic`

Pre-defined contrasts:

```{r, echo=FALSE, message=FALSE, warning=FALSE} 
apa_table(
  hypothesis2a_contrasts,
  caption = "Contrasts for the rmANOVA comparing the subjective values between $n$-back levels.",
  note = "$SE$ = standard error, $df$ = degrees of freedom, $t$ = $t$-statistic, $p$ = $p$-value, CI = confidence interval.",
  escape = FALSE,
  placement = "H"
)
```

\newpage

## Hypothesis 2b: Subjective values decline with increasing $n$-back level, even after controlling for declining task performance measured by signal detection d’ and reaction time.

Multi level model:

```{r, echo=FALSE, message=FALSE, warning=FALSE}
apa_table(h2b_result.table,
        caption = "Results of the multi level model on the influence of n-back level (as a declining logistic contrast) and task performance on subjective values.",
        escape = FALSE,
        note = "$SE$ = standard error, $df$ = degrees of freedom, $SD$ = standard deviation.",
        placement = "H")
```

The final model had an effect size of $f^{2}=$ `r round(h2b_level_f2, digits = 2)` for the $n$-back levels and $f^{2}=$ `r round(h2b_dprime_f2, digits = 2)` for $d'$.
This means that the $n$-back level explained `r paste0(formatC(h2b_level_f2 * 100, format = "f", digits = 2), "%")` and $d'$ explained `r paste0(formatC(h2b_dprime_f2 * 100, format = "f", digits = 2), "%")` of variance in SVs relative to the unexplained variance, respectively.
The beta coefficient indicated that with every 1-unit increase in $d'$, the SV increased by `r as.numeric(h2b_result.table[3,"Beta"])`.
The effect size of the median RT was $f^{2}=$ `r format(round(h2b_medianRT_f2, digits = 2), nsmall = 2)`.
The Bayes Factor of the full model against the null model was `r apa_print(h2b_BF)$statistic`.

\newpage

## Hypothesis 3a: Participants with higher NFC scores have higher subjective values for 2- and 3-back but lower subjective values for 1-back than participants with lower NFC scores.

ANOVA:

Main effect level: `r apa_print(hypothesis3a_model)$full_result$nlevels`, `r apa_print(hypothesis3a_BF_nlevel)$statistic$nlevels`

Main effect NFC group: `r apa_print(hypothesis3a_model)$full_result$nfcmedian`, `r apa_print(hypothesis3a_BF_nlevel)$statistic$nfcmedian`

Interaction: `r apa_print(hypothesis3a_model)$full_result$nfcmedian_nlevels`

Contrast for main effect level:

```{r, echo=FALSE, message=FALSE, warning=FALSE} 
apa_table(
  hypothesis3a_contrasts_nlevel,
  caption = "Paired contrast for the rmANOVA comparing the influence of Need for Cognition group and $n$-back level on difference scores of subjective values.",
  note = "The column Contrast contains the $n$ of the $n$-back levels. $SE$ = standard error, $df$ = degrees of freedom, $t$ = $t$-statistic, $p$ = $p$-value, CI = confidence interval.",
  escape = FALSE,
  placement = "H"
)
```

\newpage

```{r h3ac-plot, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Difference scores for subjective values (a) and aversiveness ratings (b) when subtracting 2- from 1-back and 3- from 2-back. Horizontal lines of the boxplots represent the median per group, whiskers represent 1.5 interquartile ranges. NFC = Need for Cognition score. $N=116$. Figure available at osf.io/vnj8x/, under a CC-BY-4.0 license.", fig.pos="H"}

  include_graphics(here("06_Paper", "COG-ED", "Stage 2", "Figures", "h3ac-plot.png"))

```

\newpage

## Hypothesis 3b: Participants with higher NFC scores have lower NASA-TLX scores in every n-back level than participants with lower NFC scores.

ANOVA:

Main effect level: `r apa_print(hypothesis3b_model)$full_result$level`, `r apa_print(hypothesis3b_BF_levels)$statistic$level`

Main effect NFC group: `r apa_print(hypothesis3b_model)$full_result$nfcmedian`, `r apa_print(hypothesis3b_BF_levels)$statistic$nfcmedian`

Interaction: `r apa_print(hypothesis3b_model)$full_result$nfcmedian_level`

Contrast for main effect level:

```{r, echo=FALSE, message=FALSE, warning=FALSE} 
apa_table(
  hypothesis3b_contrasts_levels,
  caption = "Paired contrasts for the rmANOVA comparing the influence of Need for Cognition group and $n$-back level on NASA-TLX scores.",
  note = "The column Contrast contains the $n$ of the $n$-back levels. $SE$ = standard error, $df$ = degrees of freedom, $t$ = $t$-statistic, $p$ = $p$-value, CI = confidence interval.",
  escape = FALSE,
  placement = "H"
)
```

Contrast for interaction:

```{r, echo=FALSE, message=FALSE, warning=FALSE} 
apa_table(
  hypothesis3b_contrasts_interact,
  caption = "Paired contrasts for the rmANOVA comparing the influence of Need for Cognition group and $n$-back level on NASA-TLX scores.",
  note = "The column Contrast contains the groups above (high) and below (low) the Need for Cognition score median. $SE$ = standard error, $df$ = degrees of freedom, $t$ = $t$-statistic, $p$ = $p$-value, CI = confidence interval.",
  escape = FALSE,
  placement = "H"
)
```

\newpage

```{r h3b-plot, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="NASA-TLX sum scores for each $n$-back level. Colours indicate Need for Cognition (NFC) score above or below the median. $N=116$. Figure available at osf.io/vnj8x/, under a CC-BY-4.0 license.", fig.pos="H"}

  include_graphics(here("06_Paper", "COG-ED", "Stage 2", "Figures", "h3b-plot.png"))

```

\newpage

## Hypothesis 3c: Participants with higher NFC scores have lower aversiveness ratings for 2- and 3-back but higher higher aversiveness ratings for 1-back than participants with lower NFC scores.

ANOVA:

Main effect level: `r apa_print(hypothesis3c_model)$full_result$nlevels`, `r apa_print(hypothesis3c_BF_levels)$statistic$levels`

Main effect NFC group: `r apa_print(hypothesis3c_model)$full_result$nfcmedian`, `r apa_print(hypothesis3c_BF_levels)$statistic$nfc`

Interaction: `r apa_print(hypothesis3c_model)$full_result$nfcmedian_nlevels`

Contrast for main effect level:

```{r, echo=FALSE, message=FALSE, warning=FALSE} 
apa_table(
  hypothesis3c_contrasts_levels,
  caption = "Paired contrast for the rmANOVA comparing the influence of Need for Cognition group and $n$-back level on difference scores of aversiveness ratings.",
  note = "The column Contrast contains the $n$ of the $n$-back levels. $SE$ = standard error, $df$ = degrees of freedom, $t$ = $t$-statistic, $p$ = $p$-value, CI = confidence interval.",
  escape = FALSE,
  placement = "H"
)
```

Contrast for main effect NFC group:

```{r, echo=FALSE, message=FALSE, warning=FALSE} 
apa_table(
  hypothesis3c_contrasts_nfc,
  caption = "Paired contrast for the rmANOVA comparing the influence of Need for Cognition (NFC) group and $n$-back level on difference scores of aversiveness ratings.",
  note = "$SE$ = standard error, $df$ = degrees of freedom, $t$ = $t$-statistic, $p$ = $p$-value, CI = confidence interval.",
  escape = FALSE,
  placement = "H"
)
```

\newpage

## Exploratory analysis with NFC groups by median split

ANOVA:

Main effect level: `r apa_print(explor_model)$full_result$level`, `r apa_print(explor_BF)$table[1,"statistic"]`

Main effect NFC group: `r apa_print(explor_model)$full_result$nfcmedian`, `r apa_print(explor_BF)$table[2,"statistic"]`

Interaction: `r apa_print(explor_model)$full_result$nfcmedian_level`

Contrasts for main effect level:

```{r, echo=FALSE, message=FALSE, warning=FALSE} 
apa_table(
  explor_contrasts_nlevel,
  caption = "Paired contrast for the rmANOVA comparing the influence of Need for Cognition group and $n$-back level on subjective values.",
  note = "The column Contrast contains the $n$ of the $n$-back levels. $SE$ = standard error, $df$ = degrees of freedom, $t$ = $t$-statistic, $p$ = $p$-value, CI = confidence interval.",
  escape = FALSE,
  placement = "H"
)
```

Contrasts for the interaction:

```{r, echo=FALSE, message=FALSE, warning=FALSE} 
apa_table(
  explor_contrasts_interact,
  caption = "Paired contrast for the rmANOVA comparing the influence of Need for Cognition group and $n$-back level on subjective values.",
  note = "The column Contrast contains the groups above (high) and below (low) the Need for Cognition score median. The column Level contains the $n$ of the $n$-back levels. $SE$ = standard error, $df$ = degrees of freedom, $t$ = $t$-statistic, $p$ = $p$-value, CI = confidence interval.",
  escape = FALSE,
  placement = "H"
)
```

\newpage

## Exploratory analysis with NFC as a continuous covariate

ANOVA (NFC scores standardized):

```{r, echo=FALSE, message=FALSE, warning=FALSE} 
apa_table(
  explor2_rmanova,
  caption = "Result of the rmANOVA regarding the predictive power of the $n$-back level on the subjective values when including Need for Cognition scores as a centered continuous covariate.",
  note = "NFC = Need for Cognition. $SE$ = standard error, $df$ = degrees of freedom, $F$ = $F$-statistic, $p$ = $p$-value, $\\eta_{p}^{2}$ = partial eta squared, CI = confidence interval.",
  escape = FALSE,
  placement = "H"
)
```

Estimated marginal means of the linear trends:

```{r, echo=FALSE, message=FALSE, warning=FALSE} 
apa_table(
  explor2_emm,
  caption = "Paired contrast slope analysis as a follow up for the rmANOVA regarding the predictive power of the $n$-back level on the subjective values when including Need for Cognition scores as a centered continuous covariate.",
  note = "The column $n$-back level contains the $n$ of the task levels. $SE$ = standard error, $df$ = degrees of freedom, $t$ = $t$-statistic, $p$ = $p$-value.",
  escape = FALSE,
  placement = "H"
)
```

\newpage

```{r nfcgroups-sv, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Subjective values per n-back level for participants with Need for Cognition (NFC) scores above and below the median. $N=116$. The scatter has a horizontal jitter of 0.2. Smoothing of conditional means with Loess method. Figure available at osf.io/vnj8x/, under a CC-BY-4.0 license.", fig.pos="H"}

  include_graphics(here("06_Paper", "COG-ED", "Stage 2", "Figures", "nfcgroups-sv.png"))

```